---
title: "Data Cleaning with Hotel Booking Dataset"
output: html_document
author: "Keman Xiang"
---

## Background

In this notebook, We’ll review a scenario, and focus on cleaning real data in R. We will use data cleaning functions and perform basic calculations to gain initial insights into our data.

## The scenario

In this scenario, we are junior data analysts working for a hotel booking company. We have been asked to clean a .csv file that was created after querying a database to combine two different tables from different hotels. In order to learn more about this data, We are going to need to use functions to preview the data's structure, including its columns and rows. We will also need to use basic cleaning functions to prepare this data for analysis.  


## Step 1: Load packages

Start by installing the required packages.

```{r}
#install.packages("tidyverse")
#install.packages("skimr")
#install.packages("janitor")
```

load libraries:

```{r}
library(tidyverse)
library(skimr)
library(janitor)
```
## Step 2: Import data
The data in this example is originally from the article Hotel Booking Demand Datasets (https://www.sciencedirect.com/science/article/pii/S2352340918315191), written by Nuno Antonio, Ana Almeida, and Luis Nunes for Data in Brief, Volume 22, February 2019.

The data was downloaded and cleaned by Thomas Mock and Antoine Bichat for #TidyTuesday during the week of February 11th, 2020 (https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md).

We can learn more about the dataset here:
https://www.kaggle.com/jessemostipak/hotel-booking-demand

Import data from a .csv in the project folder called "hotel_bookings.csv" and save it as a data frame called `bookings_df`:

```{r}
bookings_df <- read_csv("hotel_bookings.csv")
```
## Step 3: Getting to know our data

Before We start cleaning our data, take some time to explore it. We can use several functions to preview our data:

```{r}
head(bookings_df)
```

We can summarize or preview the data with the `str()` and `glimpse()` functions to get a better understanding of the data

```{r}
str(bookings_df)
```

```{r}
glimpse(bookings_df)
```

We can also use `colnames()` to check the names of the columns in our data set. 

```{r}
colnames(bookings_df)
```

Use the `skim_without_charts()` function from the `skimr` package:

```{r}
skim_without_charts(bookings_df)
```
## Step 4: Cleaning our data

Based on our notes we are primarily interested in the following variables: hotel, is_canceled, lead_time. Let's create a new data frame with just those columns, calling it `trimmed_df`.

```{r}
trimmed_df <- bookings_df %>% 
  select(hotel, is_canceled, lead_time)
```

Rename the variable 'hotel' to be named 'hotel_type' to be crystal clear on what the data is about:

```{r}
trimmed_df %>% 
  select(hotel, is_canceled, lead_time) %>% 
  rename(hotel_type = hotel)
```

Now, we can combine the arrival month and year into one column using the unite() function:

```{r}
comb_df <- bookings_df %>%
  select(arrival_date_year, arrival_date_month) %>% 
  unite(arrival_month_year, c("arrival_date_month", "arrival_date_year"), sep = " ")
```

## Step 5: Another way of doing things

We can also use the`mutate()` function to make changes to our columns. Let's say We wanted to create a new column that summed up all the adults, children, and babies on a reservation for the total number of people.  

```{r}
comb_df <- bookings_df %>%
  mutate(guests = adults + children + babies)

head(comb_df)
```

Great. Now it's time to calculate some summary statistics! 
Calculate the total number of canceled bookings and the average lead time for booking - We'll want to start our code after the %>% symbol. Make a column called 'number_canceled' to represent the total number of canceled bookings. Then, make a column called 'average_lead_time' to represent the average lead time.

```{r}
comb_df <- bookings_df %>%
  summarize(number_canceled = sum(is_canceled),
            average_lead_time = mean(lead_time))

head(comb_df)
```

## Activity Wrap Up
We cleaned and analyzd data in `R`; We used basic cleaning functions like `rename()` and `clean_names()` and performed basic calculations on real data. 

