---
title: "MGT 205 Project 3"
author: "KEMAN XIANG"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Project: Heart Dataset
##This project focuses on investgating, hypothesis testing ,and modeling of the heart dataset.
##Or you can also refer to my RCloud for outputs

Run the three code blocks below to get the heart data set. 


```{r}
library(tidyverse)
```


```{r}
heart <- read_csv('heart.csv')
```

```{r}
head(heart)
```

# Data Description and Usage

* **age:** The person's age in years
* **sex:** The person's sex (1 = male, 0 = female)
* **cp:** The chest pain experienced (Value 1: typical angina, Value 2: atypical angina, Value 3: non-anginal pain, Value 4: asymptomatic)
* **trestbps:** The person's resting blood pressure (mm Hg on admission to the hospital)
* **chol:** The person's cholesterol measurement in mg/dl
* **fbs:** The person's fasting blood sugar (> 120 mg/dl, 1 = true; 0 = false)
* **restecg:** Resting electrocardiographic measurement (0 = normal, 1 = having ST-T wave abnormality, 2 = showing probable or definite left ventricular hypertrophy by Estes' criteria)
* **thalach:** The person's maximum heart rate achieved
* **exang:** Exercise induced angina (1 = yes; 0 = no)
* **oldpeak:** ST depression induced by exercise relative to rest ('ST' relates to positions on the ECG plot. See more here)
* **slope:** the slope of the peak exercise ST segment (Value 1: upsloping, Value 2: flat, Value 3: downsloping)
* **ca:** The number of major vessels (0-3)
* **thal:** A blood disorder called thalassemia (3 = normal; 6 = fixed defect; 7 = reversable defect)
* **target:** Heart disease (0 = no, 1 = yes)

Input code for and answer all prompts below.


# Initial Investigation

### Question 1
#### 1a. Using the chol variable, plot the distribution as a histogram. 
```{r}
heart %>%
  ggplot(aes(x = chol)) + 
  geom_histogram()
```

#### 1b. Describe the distribution's shape (e.g. left or right skew, normal, etc).  
```{r}
#The distribution’s shape is right skewed
```




### Question 2
#### 2a. Using the age and thalach variables, find their correlation coefficient.
```{r}
cor <- cor(heart[,1],heart[,8],use='complete',method='pearson')
cor
```

#### 2b. Show this relationship in a plot. 
```{r}
heart %>%
 ggplot(aes(x=age,y=thalach)) +
 geom_point()
```

#### 2c. Does this correlation make sense for this data? Why/why not?
```{r}
#This correlation makes sense. Because as age increases, human’s physical functions and health reduce, therefore have lower heart rates.
```



# Hypothesis Testing

### Question 3
#### 3a. Perform a t test for difference in means between genders (sex) on chol. 
```{r}
t.test(chol ~ sex, data = heart)
```

#### 3b. Find and show the mean cholesterol of men and women.
```{r}
heart %>%
 group_by(sex) %>%
 summarise(mean_cholesterol = mean(chol))
```

#### 3c. Interpret the results above using a significance threshold of 0.05. Are the results statistically significant? 
```{r}
#Because the p-value = 0.002985 < significance threshold 0.05, the results are statistically significant.
```


### Question 4
#### 4a. Perform a t test (hypothesis test) to determine if the blood pressure among those with heart disease is higher or not. 
```{r}
t.test(trestbps ~ target, data = heart, alternative = 'less')
```

#### 4b. Interpret these results. Statistically, is blood pressure higher among those with heart disease? How can you tell? 
```{r}
#The results are not statistically significant. We fail to reject the null hypothesis. There is insufficient evidence to conclude that resting blood pressure is higher among individuals with heart disease.
```



# Modelling

### Question 5
#### 5a. Separate heart data into a training and testing data set.
```{r}
sample_indexes <- sample(1:nrow(heart), size = nrow(heart) * 3/4)
heart_train <- heart[sample_indexes, ]
heart_test <- heart[-sample_indexes, ]
```

#### 5b. Construct a LINEAR regression model to predict for blood pressure. 
```{r}
fit <- lm(trestbps ~ ., data = heart_train)
summary(fit)
```

#### 5c. What are the statistically significant variables in this model? What is the value of the Multiple R-squared statistic, and what does this value mean? Overall, is this a good prediction model? 
```{r}
#1. Statistically significant variables: age (because p_value 0.000134 < 0.05) fbs (because p_value 0.015681 < 0.05) oldpeak (because p_value 0.044594 < 0.05)
#2. Multiple R-squared: 0.1531 R^2=1-SSE/SST (SSE:error sum of squares, SST:total sum of squares) 
  #It refers to the goodness of fit, which is how well the regression line fits the observations. Its value is between 0 and 1, the closer to 1, the better the regression fitting effect. Because Multiple R-squared is 0.1531, the regression fits not well and is not a good prediction model.
```



### Question 6
#### 6a. Construct a LOGISTIC regression model to predict for heart disease indicator flag. Calculate the accuracy using the testing data.
```{r}
fit_class <- glm(target ~ ., data = heart_train, family = binomial(link = 'logit'))

prediction_class <- ifelse(predict(fit_class, newdata = heart_test, type = 'response') < 0.5, 0, 1)

sum(heart_test$target == prediction_class)/length(prediction_class)
```

#### 6b. What percent of the time is this model accurate?
```{r}
#81.57895% of the time this model is accurate
```

